\huge \systemEnding: Growing Libraries of Concepts with Wake-Sleep Program Induction

### 2018-07-22 misc2.tex
improvements to poster

\title{\huge \systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning} % Poster title

\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)

@@

\begin{alertblock}{Wake/Sleep DSL Induction}
  \flushleft
  \textbf{Domain Specific Language (DSL):}
  A finely-tuned program representation, specialized to a domain of programming tasks.
  Prior work in program learning  largely uses hand-engineered DSLs.

  \textbf{Approach:}  \system algorithm,
  which bootstraps a learned DSL while jointly training
  a neural net to search for programs in the learned DSL.
Given a few hundred programming tasks, alternatingly:
  \begin{itemize}
  \item \textbf{Wake}: synthesize programs
  \item \textbf{Sleep-R}: train neural net (\textbf{R}ecognition model)
    \item \textbf{Sleep-G}: improve DSL (\textbf{G}enerative model)
    \end{itemize}

@@

\begin{tikzpicture}[scale=4,line width=1.5mm]

  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
  \plate {}{(zp)(xp)}{};
%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
%  \draw [->,red] (tx.west) -- (zp.east);
  \draw [->] (dx.east) -- (zp.west);
  \draw [->] (zp.east) -- (xp.west);

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };

@@

  Observe $N$ \textbf{tasks}, written $\left\{x_n \right\}_{n = 1}^N$, each a program synthesis problem.

  Solve task $x_n$ with latent program $p_n$

  \textbf{Likelihood model} $\probability[x_n|p_n]$
  scores program $p_n$ on task $x_n$

  Latent  \textbf{DSL $\mathcal{D}$} acts as generative model over programs: $\probability[x|\mathcal{D}]$

$$
\underbrace{p_n^* =   \argmax_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}^*]}_{\text{\textbf{Wake}}}
$$
\vspace{1cm}
$$
\underbrace{\mathcal{D}^* = \argmax_{\mathcal{D}}    \probability[\mathcal{D}]\prod_n\sum_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}]}_{\textbf{\text{Sleep-G}}}
$$

@@

\begin{block}{Neural recognition model}

  Neural network $Q(p|x)$
  predicts distribution over programs conditioned on tasks.
  Simple $Q$: just predicts probabilities of DSL productions.  Goal: learn to invert generative model

$$\underbrace{\min_Q \text{KL}\left(\probability[p|x,\mathcal{D}] || Q(p|x)\right) }_{\text{\textbf{Sleep-R}}}$$

@@

  \begin{itemize}
  \item \textbf{Samples (``Dreams'') from DSL}: Unlimited data, but only high-quality if generative model $\mathcal{D}$ is good. Like Helmholtz Machine's recognition model training. Loss:
    \begin{equation*}
\expect_{(p,x)\sim\mathcal{D} }\left[\log Q(p|x)\right]
      \end{equation*}
  \item \textbf{Self-Supervised}: $(x_n,p_n)$ pairs discovered during waking. Loss:
    \begin{equation*}
 %\sum_{n = 1}^N\left[\sum_{(x_n,p_n)}
  \frac{\probability\left[x_n,p_n|\mathcal{D}\right]}{\sum_{(x_n,p_n')}\probability\left[x_n,p_n|\mathcal{D} \right]}\log Q(p_n|x_n)%\right]
      \end{equation*}
  \end{itemize}

@@

\begin{block}{Fragment Grammars: Inducing a DSL}
Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)
\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{7.5cm}
  \begin{tikzpicture}[scale=2,line width=1mm]
    \node [rotate=90] at (-2.5,-0.5) {prog.};

>>>>

\title{\huge \systemEnding: Growing Libraries of Concepts with Wake-Sleep Program Induction} % Poster title

\author{Kevin Ellis, Mathias Sabl\'e Meyer, Lucas Morales, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)

@@

\begin{alertblock}{Wake/Sleep DSL Induction}
  \flushleft
  Goal: acquire domain-specific knowledge needed to induce a class of programs
  \begin{itemize}
  \item Declarative knowledge: library of concepts, `DSL' (domain specific language)
    \item Procedural knowledge: knowing how and when to use those concepts; search strategy
    \end{itemize}

%%   \system algorithm,
%%   which bootstraps a learned DSL while jointly training
%%   a neural net to search for programs in the learned DSL.
%% Given a few hundred programming tasks, alternatingly:
  \begin{itemize}
  \item \textbf{Wake}: synthesize programs
  \item \textbf{Sleep-R}: train neural net (\textbf{R}ecognition model)
    \item \textbf{Sleep-G}: improve DSL (\textbf{G}enerative model)
    \end{itemize}

@@

    \includegraphics[width = 20cm]{figures/animation/DC.eps}

%% \begin{tikzpicture}[scale=4,line width=1.5mm]

%%   \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
%%   \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
%%   \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
%%   \plate {}{(zp)(xp)}{};
%% %  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
%%   \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
%% %  \draw [->,red] (tx.west) -- (zp.east);
%%   \draw [->] (dx.east) -- (zp.west);
%%   \draw [->] (zp.east) -- (xp.west);

%%   \node at (nn) {
%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
%%       \node[rectangle] at (1,1) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (I-\name) at (0,-\y) {};
%%       \foreach \name / \y in {1,...,3}
%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (O-\name) at (2,-\y) {};
%%       \foreach \source in {1,...,4}
%%           \foreach \dest in {1,...,3}
%%               \draw [-latex] (I-\source) -- (H-\dest);
%%       \foreach \source in {1,...,3}
%%           \foreach \dest in {1,...,4}
%%               \draw [-latex] (H-\source) -- (O-\dest);
%%     \end{tikzpicture}
%%   };
%%   \node[shift={+(0,-1.7)}] at (nn) { $Q$  };

%%   \end{tikzpicture}

  Observe \textbf{tasks} (program induction problems). 
  Solve tasks w/ latent programs.
  %% \textbf{Likelihood model} $\probability[x_n|p_n]$
  %% scores program $p_n$ on task $x_n$
  Latent  \textbf{DSL} acts as generative model over programs: `library of concepts'.
  \textbf{Recognition model}: a neural net that predicts distribution over programs conditioned on task: `procedural knowledge'

\begin{block}{Wake: Problem solving}
  \begin{tikzpicture}[scale=2.5]
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.4,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.6cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
%    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};

    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node[align=left] at (7,-1) (f) {Programs};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
%    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};

 %   \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};

@@

@@

\begin{block}{Sleep-R: Procedural knowledge}

  Neural network
  predicts distribution over programs conditioned on tasks.
  Simple model: just predicts probabilities of DSL productions.  Goal: learn to invert generative model from (program, task) pairs

@@

  \begin{itemize}
  \item \textbf{``Dreaming'' from DSL}:

    \includegraphics[width = 20cm]{figures/animation/Sleep-R-2.eps}
  \item \textbf{Experience replay}: (program, task) pairs discovered during waking.
  \includegraphics[width = 20cm]{figures/animation/Sleep-R-1.eps}    
  \end{itemize}

@@

\begin{block}{Sleep-G: Growing a DSL}
Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)
\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{7.5cm}
  \begin{tikzpicture}[scale=2,line width=1mm]
    \node [rotate=90] at (-2.5,-0.5) {prog.};
